---
title: "Homework 2"
author: "PSTAT 131/231"
output:
    html_document:
      toc: true
      toc_float: true
      code_folding: show
      
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = FALSE)
```

## Linear Regression

```{r}
library(readr)
library(magrittr)
abalone<-read_csv('abalone.csv')
```

### Question 1

Your goal is to predict abalone age, which is calculated as the number of rings plus 1.5. Notice there currently is no `age` variable in the data set. Add `age` to the data set.

Assess and describe the distribution of `age`.

```{r}
abalone$age<-abalone$rings+1.5
plot(abalone$age)
boxplot(abalone$age)
```
The most of abalone are about 10 years old, and a few reach 30.

### Question 2

Split the abalone data into a training set and a testing set. Use stratified sampling. You should decide on appropriate percentages for splitting the data.

*Remember that you'll need to set a seed at the beginning of the document to reproduce your results.*
```{r}
set.seed(123)
abalone_splitting<-sample(c(TRUE, FALSE), nrow(abalone), replace=TRUE, prob=c(0.7,0.3))
training_set<-abalone[abalone_splitting,]
test_set<-abalone[!abalone_splitting,]
dim(abalone)
dim(training_set)
dim(test_set)
```
### Question 3

Because the relationship between rings and age is only 1.5 difference, not proportional, we cannot use rings to predict age.


```{r}
library(recipes)
abalone_recipe<-recipe(age~ ., data=training_set) 
step_dummy(abalone_recipe,all_nominal_predictors())
step_interact(abalone_recipe,terms=~type:shucked_weight)
step_interact(abalone_recipe,terms=~longest_shell:diameter) 
step_interact(abalone_recipe,terms=~shucked_weight:shell_weight)
step_center(abalone_recipe,all_predictors())
step_scale(abalone_recipe,all_predictors())
```


### Question 4

Create and store a linear regression object using the `"lm"` engine.
```{r}
library(parsnip)
lm_model <- linear_reg() %>% 
  set_engine("lm")
```

### Question 5

```{r}
library(workflows)
library(workflowsets)
lm_wflow <- workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(abalone_recipe)
```

### Question 6

```{r}
library(generics)
lm_fit <- fit(lm_wflow, training_set)
extract_fit_parsnip(lm_fit)
```


### Question 7

```{r}
train_res <- predict(lm_fit, new_data=training_set %>% select(-age))
train_res %>%
  head()
```

```{r}
train_res<- bind_cols(train_res, training_set %>% select(age))
train_res %>%
  head()
```

```{r}
library(yardstick)
rmse(train_res, truth = age, estimate = .pred)
metrics <- metric_set(rmse, rsq, mae)
metrics(train_res, truth = age, estimate = .pred)
```


